---
title: "Lab4_Demo"
author: "Mateo Robbins"
date: "2024-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)     # for data wrangling
library(ggplot2)   # for awesome plotting
library(rsample)   # for data splitting
library(recipes)   # data preprocessing
library(caret)     # for logistic regression modeling
```
Let's explore how employee income and overtime hours worked affect likelihood of employee attrition.  Any predictions?

```{r}
#
data("attrition", package = "modeldata")

df <- attrition %>% mutate_if(is.ordered, factor, ordered = FALSE)

# Create training (70%) and test (30%) sets for the 
# rsample::attrition data.
set.seed(123)  # for reproducibility 
churn_split <- initial_split(attrition, prop = .70)
churn_train <- training(churn_split)
churn_test  <- testing(churn_split)
```

Let's set up set up our recipes now for preprocessing. 
```{r recipe}
#specify and prep recipe
churn_rec <- recipe(Attrition ~., data = churn_train) %>% 
  step_integer(Attrition, zero_based = T) %>% 
  prep(churn_train)

#bake recipe with training data
churn_baked_train <- bake(churn_rec, new_data = churn_train)
```

```{r specify_models_glm}
#MonthlyIncome
model_inc <- glm(Attrition ~ MonthlyIncome, family = "binomial", data = churn_baked_train)
  
#OverTime
model_time <- glm(Attrition ~ OverTime, family = "binomial", data = churn_baked_train)
```


```{r tidy_model_objs}
broom::tidy(model_inc)
broom::tidy(model_time)
```



```{r exp_coefs}
#exponentiate the coefficients from model objects for interpretation. Gives us changes in odds of attrition
exp(coef(model_inc))
exp(coef(model_time))
```

for each dollar, odds of them leaving ( prob of them leaving / prob of not leaving) is incrementally decreasing. multiplity by .99 ( so its decreasing). odd of them leaving multipled by 3.5 times if they worked overtime. (change in the odds of the event occuring , multiplicateively) when an employee works overtime, the odds of them leaving are 3 times more likely. 

```{r recode_attrition_test}
churn_baked_test <-  bake(churn_rec, new_data = churn_test)
```

```{r plot_income_attrition}
ggplot(churn_baked_test, aes(x = MonthlyIncome, y = Attrition))+
  geom_point()+
  stat_smooth(method = "glm", se = T, method.args  = list(family = "binomial"))
```

We can add more predictors, creating a multiple logistic regression model

```{r mult_log_regression}

model_both <- glm(
  Attrition ~ MonthlyIncome + OverTime,
  family = "binomial", 
  data = churn_baked_train
  )
```


```{r}
ggplot(churn_baked_test, aes(x = MonthlyIncome, y = Attrition, color = OverTime))+
  geom_point(alpha = .3)+
  stat_smooth(method = "glm", se = F, method.args = list(family = "binomial "))
```

